---
output:
  pdf_document: default
  html_document: default
---
# Analysis of the Triboro line

## Requirements

### Libraries
```{r, message=FALSE}
library(tidyverse)
```

### Files
```{r, message=FALSE}
subway_lines <- readr::read_csv('data/nta-subway-lines.csv')
subway_times <- readr::read_csv('data/nta-subway-times.csv')
driving_times <- readr::read_csv('data/nta-driving-times.csv')
walking_times <- readr::read_csv('data/nta-walking-times.csv')
```

### Data
```{r}

subway_lines <- subway_lines %>%
  dplyr::mutate(
    log_S000 = log(S000)
  )
```

## Exploratory data analysis

### Distribution of job counts
Original

Natural log  

### Distribution of commute counts
Original
```{r, message = FALSE}
commute_counts <- subway_lines %>%
  dplyr::select(trip, S000) %>%
  dplyr::mutate(
    log_S000 = log(S000)
  )
```


```{r}
ggplot(commute_counts) +
  geom_histogram(aes(x = S000), fill = "steelblue", color = "grey", binwidth = 30)
```

Natural log
```{r}
ggplot(data = commute_counts) +
  geom_histogram(aes(x = log_S000), bins = 30, fill = "steelblue", color = "grey")
```

### Number of subway lines and commute count
Original

Transformed

### Subway Transit time and commute count
Original

Transformed

### Driving in traffic time and commute count
Original

Transformed

### Walking time and commute count

Original

Transformed

## Regression of Subway, Driving, and walking

### Subway model

### Driving

### Walking

### Multiple linear regression for all three factors

### Equations plotted for all factors
Along all axis

Cut to most pivotal times (10 to 50 minutes)
Table of values at 10, 25, 50

## Auto Correlation of Subway, Driving, and Walking


### Global Moran's I

#### Subway
Statistic
Plot 

#### Driving
Statisic
Plot 

#### Walking
Statustic
Plot

### LISA

#### Driving
Statistic 
Plot with major roadways overlayed

#### Walking
Statistic
Plot of local area

#### Subway
Statistic 
Plot with subway lines

#### Network autocorrelation

#### Visualization of network

#### Visualization of network's complement

#### Global Moran's I

#### LISA
Plot by coloring desire lines