---
output:
  pdf_document: default
  html_document: default
---
# Analysis of the Triboro line

## Requirements

### Libraries
```{r, message=FALSE}
library(tidyverse)
library(igraph)
library(sf)
```

### Files
```{r, message=FALSE}
nys_od <- readr::read_csv('data/ny_od_main_JT00_2019.csv')
nyc_nta_borders <- sf::st_read('data/nyc_2010_nta_borders.geojson')
nyc_nta_tract_equiv <- readxl::read_xlsx('data/nyc_2010_census_tract_nta_equiv.xlsx')

subway_lines <- readr::read_csv('data/nta-subway-lines.csv')
subway_times <- readr::read_csv('data/nta-subway-times.csv')
driving_times <- readr::read_csv('data/nta-driving-times.csv')
walking_times <- readr::read_csv('data/nta-walking-times.csv')
```

### Data
```{r, message=FALSE}
bk_name <- "Brooklyn"
bk_county_code <- "047"
bk_parks <- "BK99"

bk_nta_border <- nyc_nta_borders %>%
  dplyr::filter(BoroName == bk_name)  %>%
  dplyr::filter(NTACode != bk_parks) %>%
  dplyr::select("NTACode")

bk_nta_tract_equiv <- nyc_nta_tract_equiv %>%
  dplyr::filter(borough_name == bk_name) %>%
  dplyr::filter(nta_code != bk_parks) %>%
  dplyr::rename(tract = census_tract) %>%
  dplyr::select("tract", "nta_code")

od <- nys_od %>%
  dplyr::filter(
    stringr::str_sub(as.character(w_geocode), 3, 5) == bk_county_code &
    stringr::str_sub(as.character(h_geocode), 3, 5) == bk_county_code 
  ) %>%
  dplyr::mutate(
   w_tract = stringr::str_sub(as.character(w_geocode), 6, 11)
  ) %>%
  dplyr::mutate(
   h_tract = stringr::str_sub(as.character(h_geocode), 6, 11)
  ) %>%
  dplyr::select("w_tract", "h_tract", "S000") %>%
  dplyr::left_join(bk_nta_tract_equiv, c("w_tract" = "tract")) %>%
  dplyr::rename(w_nta_code = nta_code) %>%
  dplyr::left_join(bk_nta_tract_equiv, c("h_tract" = "tract")) %>%
  dplyr::rename(h_nta_code = nta_code) %>%
  dplyr::select("h_nta_code", "w_nta_code", "S000") %>%
  dplyr::filter(w_nta_code != bk_parks & h_nta_code != bk_parks) 
``` 

## Exploratory data analysis

### Distribution of job counts
```{r, message = FALSE}
job_counts <- od %>%
  dplyr::group_by(w_nta_code) %>%
  dplyr::summarise(
    S000 = sum(S000)
  ) %>% 
  unique() %>%
  mutate(
    log_S000 = log(S000)
  )
```

Original
```{r, message=FALSE}
ggplot(job_counts) +
  geom_histogram(aes(x = S000), fill = "steelblue", color = "grey", bins = "30")
```

Natural log  
```{r, message=FALSE}
ggplot(job_counts) +
  geom_histogram(aes(x = log_S000), fill = "steelblue", color = "grey", bins = "30")
```

### Distribution of commute counts
```{r, message = FALSE}
commute_counts <- subway_lines %>%
  dplyr::select(trip, S000) %>%
  dplyr::mutate(
    log_S000 = log(S000)
  )
```

Original
```{r}
ggplot(commute_counts) +
  geom_histogram(aes(x = S000), fill = "steelblue", color = "grey", binwidth = 30)
```

Natural log
```{r}
ggplot(data = commute_counts) +
  geom_histogram(aes(x = log_S000), bins = 30, fill = "steelblue", color = "grey")
```

### Number of subway lines and commute count
```{r, message = FALSE}
subway_lines <- subway_lines %>%
  dplyr::mutate(
    line_count_ordinal = as.character(line_count),
    log_S000 = log(S000)
  )
```
Original
```{r}
ggplot(data = subway_lines, aes(x = line_count_ordinal, y = S000)) +
  geom_boxplot()
```

Transformed
```{r, message = FALSE}
ggplot(data = subway_lines, aes(x  = line_count_ordinal, y = log_S000)) +
  geom_boxplot()
```

### Subway Transit time and commute count
```{r, message=FALSE}
subway_times <- subway_times %>%
  dplyr::mutate(
    log_S000 = log(S000),
    i_seconds_in_transit = 1 / seconds_in_transit^(1/8)
  )
subway_times_connected <- subway_times %>%
  dplyr::filter(
    subway_lines$line_count > 0 
  )
```

Original
```{r, message = FALSE}
ggplot(data = subway_times_connected, aes(x = seconds_in_transit, y = S000)) +
  geom_point() +
  stat_smooth()
```

Transformed
```{r, message = FALSE}
ggplot(data = subway_times_connected, aes(x = i_seconds_in_transit, y = log_S000)) +
  geom_point() +
  stat_smooth()
```


### Driving in traffic time and commute count
```{r, message=FALSE}
driving_times <- driving_times %>%
  dplyr::mutate(
    log_S000 = log(S000),
    i_seconds_in_traffic = 1 / seconds_in_traffic^(1/8)
  )
```
Original
```{r, message = FALSE}
ggplot(data = driving_times, aes(x = seconds_in_traffic, y = S000)) +
  geom_point() +
  stat_smooth()
```

Transformed
```{r, message = FALSE}
ggplot(data = driving_times, aes(x = i_seconds_in_traffic, y = log_S000)) +
  geom_point() +
  stat_smooth()
```

### Walking time and commute count
```{r, message=FALSE}
walking_times <- walking_times %>%
  dplyr::mutate(
    log_S000 = log(S000),
    i_seconds_of_walking = 1 / seconds_of_walking^(1/8)
  )
```

Original
```{r, message = FALSE}
ggplot(data = walking_times, aes(x = seconds_of_walking, y = S000)) +
  geom_point() +
  stat_smooth()
```

Transformed
```{r, message = FALSE}
ggplot(data = walking_times, aes(x = i_seconds_of_walking, y = log_S000)) +
  geom_point() +
  stat_smooth()
```

## Regression of Subway, Driving, and walking

### Subway model
```{r, message=FALSE}
subway_connected_model <- lm(subway_times_connected$log_S000 ~ subway_times_connected$i_seconds_in_transit)
summary(subway_connected_model)
plot(subway_connected_model)
```

### Driving
```{r, message=FALSE}
driving_model <- lm(driving_times$log_S000 ~ driving_times$i_seconds_in_traffic)
summary(driving_model)
plot(driving_model)
```

### Walking
```{r, message = FALSE}
walking_model <- lm(walking_times$log_S000 ~ walking_times$i_seconds_of_walking)
summary(walking_model)
plot(driving_model)
```

### Multiple linear regression for all three factors

### Equations plotted for all factors
```{r, message=FALSE}
subway_connected_eq <- function(t) exp(-3.0033 + 22.1264 / t^(1/8))
driving_eq <- function(t) exp(-1.9856 + 17.9177 / t^(1 / 8))
walking_eq <- function(t) exp(-1.9091 + 20.8372 / t^(1 / 8))

summary(subway_times_connected$seconds_in_transit)
summary(driving_times$seconds_in_traffic)
summary(walking_times$seconds_of_walking)
```

```{r, message = FALSE}
ggplot(
  dplyr::data_frame(
    seconds = seq(from = 301, to = 15200, by = 14.9)
  ), aes(seconds)) +
  stat_function(fun = subway_connected_eq, color = "steelblue", xlim = c(645, 6924)) +
  stat_function(fun = driving_eq, color = "seagreen", xlim = c(276, 3122)) +
  stat_function(fun = walking_eq, color = "goldenrod", xlim = c(702, 15171))
```


Along all axis

Cut to most pivotal times (10 to 50 minutes)
Table of values at 10, 25, 50
```{r, message = FALSE}
ggplot(
  dplyr::data_frame(
    seconds = seq(from = 720, to = 3000, by = 2.28)
  ), aes(seconds)) +
  stat_function(fun = subway_connected_eq, color = "steelblue") +
  stat_function(fun = driving_eq, color = "seagreen") +
  stat_function(fun = walking_eq, color = "goldenrod")
```
## Auto Correlation of Subway, Driving, and Walking


### Global Moran's I
```{r, message = FALSE}
subway_times <- subway_times %>%
  dplyr::mutate(
    i_c_seconds_in_transit = ifelse(subway_lines$line_count > 0, i_seconds_in_transit, 0)
  ) 

subway_graph <- subway_times %>%
  dplyr::select(
    c(
      nta_one,
      nta_two,
      i_c_seconds_in_transit
    )
  ) %>%
  dplyr::rename(
    from = nta_one,
    to = nta_two,
    weight = i_c_seconds_in_transit,
  ) %>%
  igraph::graph.data.frame(
    directed = FALSE
  ) 

subway_weights <- subway_graph %>%
  igraph::as_adjacency_matrix(attr = "weight") %>%
  spdep::mat2listw()

driving_weights <- driving_times %>%
  dplyr::select(
    c(
      nta_one,
      nta_two,
      i_seconds_in_traffic
    )
  ) %>%
  dplyr::rename(
    from = nta_one,
    to = nta_two,
    weight = i_seconds_in_traffic
  ) %>%
  igraph::graph.data.frame(
    directed = FALSE
  ) %>%
  igraph::as_adjacency_matrix(attr = "weight") %>%
  spdep::mat2listw()

walking_weights <- walking_times %>%
  dplyr::select(
    c(
      nta_one,
      nta_two,
      i_seconds_of_walking
    )
  ) %>%
  dplyr::rename(
    from = nta_one,
    to = nta_two,
    weight = i_seconds_of_walking
  ) %>%
  igraph::graph.data.frame(
    directed = FALSE
  ) %>%
  igraph::as_adjacency_matrix(attr = "weight") %>%
  spdep::mat2listw()
```

#### Subway
```{r, message=FALSE}
subway_global_morans <- spdep::moran.test(
  job_counts$log_S000,
  subway_weights,
  zero.policy = TRUE,
)
print(subway_global_morans)

spdep::moran.plot(
  job_counts$log_S000,
  subway_weights,
  zero.policy = TRUE,
  xlab = "Job count",
  ylab = "Lagged job count"
)
```
#### Driving
```{r, message=FALSE}
driving_global_morans <- spdep::moran.test(
  job_counts$log_S000,
  driving_weights,
  zero.policy = TRUE,
)
print(driving_global_morans)

spdep::moran.plot(
  job_counts$log_S000,
  driving_weights,
  zero.policy = TRUE,
  xlab = "Job count",
  ylab = "Lagged job count"
)
```

#### Walking
```{r, message=FALSE}
walking_global_morans <- spdep::moran.test(
  job_counts$log_S000,
  walking_weights,
  zero.policy = TRUE,
)
print(walking_global_morans)

spdep::moran.plot(
  job_counts$log_S000,
  walking_weights,
  zero.policy = TRUE,
  xlab = "Job count",
  ylab = "Lagged job count"
)
```

### LISA
```{r, message=FALSE}
avg_jobs <- mean(job_counts$log_S000)

classify_co_types <- function(mode_lisa, l_job_counts, avg_job_count) {
  mode_lisa %>%
    tibble::as_tibble() %>%
    magrittr::set_colnames(
      c("Ii","E.Ii","Var.Ii","Z.Ii","Pr(z > 0)") 
    ) %>%
    dplyr::mutate(
      co_type = dplyr::case_when(
        `Pr(z > 0)` <= 0.05 &
          Ii >= 0 &
          l_job_counts >= avg_job_count ~ "HH",
        `Pr(z > 0)` <= 0.05 &
          Ii >= 0 &
          l_job_counts <  avg_job_count ~ "LL",
        `Pr(z > 0)` <= 0.05 &
          Ii < 0 &
          l_job_counts >= avg_job_count ~ "HL",
        `Pr(z > 0)` <= 0.05 &
          Ii < 0 &
          l_job_counts < avg_job_count ~ "LH"
      )
    )
}

subway_lisa <- spdep::localmoran(
  job_counts$log_S000,
  subway_weights,
  zero.policy = TRUE,
  na.action = na.omit
) 

driving_lisa <- spdep::localmoran(
  job_counts$log_S000,
  driving_weights,
  zero.policy = TRUE,
  na.action = na.omit
)

walking_lisa <- spdep::localmoran(
  job_counts$log_S000,
  walking_weights,
  zero.policy = TRUE,
  na.action = na.omit
)

subway_classes <- classify_co_types(subway_lisa, job_counts$log_S000, avg_jobs)
driving_classes <- classify_co_types(driving_lisa, job_counts$log_S000, avg_jobs)
walking_classes <- classify_co_types(walking_lisa, job_counts$log_S000, avg_jobs)

subway_bk_nta_border <- bk_nta_border %>%
  dplyr::mutate(
    co_type = ifelse(is.na(subway_classes$co_type), "Insignificant", subway_classes$co_type)
  )

driving_bk_nta_border <- bk_nta_border %>%
  dplyr::mutate(
    co_type = ifelse(is.na(driving_classes$co_type), "Insignificant",  driving_classes$co_type)
  )

walking_bk_nta_border <- bk_nta_border %>%
  dplyr::mutate(
    co_type = ifelse(is.na(walking_classes$co_type), "Insignificant", walking_classes$co_type)
  )
```


#### Subway
Statistic 
Plot with subway lines
```{r, message=FALSE}
ggplot(subway_bk_nta_border) +
  geom_sf(aes(fill = co_type), col = 'lightgrey') +
  scale_fill_manual(
    values = c("red", "goldenrod", "NA", "steelblue"),
    name = "Clusters & \nOutliers"
  ) +
  labs(
    title = "Neighboring job counts"
  )
```
#### Driving

```{r, message=FALSE}
ggplot(driving_bk_nta_border) +
  geom_sf(aes(fill = co_type), col = 'lightgrey') +
  scale_fill_manual(
    values = c("red", "goldenrod", "NA", "steelblue", "seagreen"),
    name = "Clusters & \nOutliers"
  ) +
  labs(
    title = "Neighboring job counts"
  )
```
Statistic 
Plot with major roadways overlayed

#### Walking
```{r, message=FALSE}
ggplot(walking_bk_nta_border) +
  geom_sf(aes(fill = co_type), col = 'lightgrey') +
  scale_fill_manual(
    values = c("red", "goldenrod", "NA", "seagreen"),
    name = "Clusters & \nOutliers"
  ) +
  labs(
    title = "Neighboring job counts"
  )
```
subway
Statistic
Plot of local area
#### Network autocorrelation

#### Visualization of network

#### Visualization of network's complement

#### Global Moran's I

#### LISA
Plot by coloring desire lines