---
output:
  pdf_document: default
  html_document: default
---
# Network auto-correlation for desire lines between NTAs in Brooklyn

Building off the network of the desire-lines from the previous section, the graph
is transformed to describe the relationship between the desire lines. After this transformation,
a node is defined as a trip. The nodes are named for the two NTAs that are part of
the trip. Because the base graph is undirected, the order of the codes in the trip
name are not related to the direction of the trips.
An edge exists between two trips if the trips have an
NTA in common. This graph is also undirected.

Network auto-correlation is performed with the graph in this form. The value of
each node is the number of commutes that are made along that trip. 

```{r, message=FALSE}
library(tidyverse)
library(tidygraph)
library(ggraph)
library(spData)
library(spdep)
library(igraph)
```

Define each trip as a node
```{r, message=FALSE}
nta_trips <- readr::read_csv('./data/nta-trip-network.csv')
nta_trip_nodes <- tibble::tibble(name = nta_trips$trip, trip_count = nta_trips$S000)
```

Define a utility function that builds the edges between trips
```{r, message=FALSE}
build_edges <- function(nodes){
  edges_from <- vector()
  edges_to <- vector()
  nodes_count <- length(nodes)
  for(i in 1:(nodes_count - 1)) {
    offset <- i + 1
    from_node <- nodes[i]
    from_nta_one <- stringr::str_sub(from_node, 1, 4)
    from_nta_two <- stringr::str_sub(from_node, 5, 8)
    for(j in offset:nodes_count){
      to_node <- nodes[j]
      are_neighbors <- stringr::str_detect(to_node, from_nta_one) | stringr::str_detect(to_node, from_nta_two)
      if (are_neighbors) {
        edges_from <- append(edges_from, from_node)
        edges_to <- append(edges_to, to_node)
      }
    }
  }
  return (tibble::tibble(from = edges_from, to = edges_to))
}
```

## Distribution of trips
Orginal data
```{r}
ggplot(nta_trip_nodes, aes(x = trip_count)) +
  geom_histogram()
```

Transform data with natural log
```{r}
ggplot(nta_trip_nodes, aes(x = log(trip_count))) +
  geom_histogram()
```

log base 10
```{r}
ggplot(nta_trip_nodes, aes(x = log10(trip_count))) +
  geom_histogram()
```

Construct the graph
```{r, message=FALSE}
nta_trip_edges <- build_edges(nta_trip_nodes$name) 
nta_trip_network <- tidygraph::tbl_graph(nodes = nta_trip_nodes, edges = nta_trip_edges)
```

## Subgraph
Use a sub section of the network to make it easier to visualize

### Random sample
Construct a subgraph with a random sample of nodes
```{r, message=FALSE}
nta_trip_nodes_rand <- nta_trip_nodes %>%
  dplyr::slice_sample(n = 50)
nta_trip_edges_rand <- build_edges(nta_trip_nodes_rand$name)
nta_trip_network_rand <- tidygraph::tbl_graph(nodes = nta_trip_nodes_rand, edges = nta_trip_edges_rand)

total_trips_rand <- sum(nta_trip_nodes_rand$trip_count)
ggraph::ggraph(nta_trip_network_rand, layout="stress") +
  geom_edge_link() +
  geom_node_circle(aes(r = (nta_trip_nodes_rand$trip_count / total_trips_rand)), fill = "blue") +
  geom_node_text(aes(label = stringr::str_c(stringr::str_sub(name, 3,4), '&', stringr::str_sub(name, 7,8))), repel = TRUE)
```

### Most popular trips
Construct a subgraph using the nodes with the top 50 trip counts
```{r, message=FALSE}
nta_trip_nodes_top <- nta_trip_nodes %>%
  dplyr::slice_max(order_by = trip_count, n = 50)
nta_trip_edges_top <- build_edges(nta_trip_nodes_top$name)
nta_trip_network_top <- tidygraph::tbl_graph(nodes = nta_trip_nodes_top, edges = nta_trip_edges_top)

total_trips_top <- sum(nta_trip_nodes_top$trip_count)
ggraph::ggraph(nta_trip_network_top, layout="stress") +
  geom_edge_link() +
  geom_node_circle(aes(r = (nta_trip_nodes_top$trip_count / total_trips_top)), fill = "blue") +
  geom_node_text(aes(label = stringr::str_c(stringr::str_sub(name, 3,4), '&', stringr::str_sub(name, 7,8))), repel = TRUE)
```

## Global Moran's I
Calculate the total Global Moran's I for the whole network.
```{r, message=FALSE}
nta_trip_network_weights <- nta_trip_network %>%
  igraph::as_adj() %>%
  spdep::mat2listw()

global_morans <- spdep::moran.test(nta_trip_nodes$trip_count, nta_trip_network_weights, zero.policy = TRUE)
global_morans  
```
Plot the Global Moran's I
```{r, message=FALSE}
spdep::moran.plot(
  nta_trip_nodes$trip_count,
  nta_trip_network_weights,
  zero.policy = TRUE,
  xlab = "trip count",
  ylab = "lagged trip count",
  pch = 20,
)
```

### Natural log for trips
```{r, message=FALSE}
t_global_morans <- spdep::moran.test(log(nta_trip_nodes$trip_count), nta_trip_network_weights, zero.policy = TRUE)
t_global_morans 
```

## Local Indicators of spatial autocorrelation
```{r, message=FALSE}
local_moran <- spdep::localmoran(
  nta_trip_nodes$trip_count,
  nta_trip_network_weights,
  zero.policy = TRUE,
  na.action = na.omit,
)

sig_lev <- 0.05
avg_trip_count <- mean(nta_trip_nodes$trip_count)

lisa_classes <- local_moran %>%
  tibble::as_tibble() %>%
  magrittr::set_colnames(
    c("Ii","E.Ii","Var.Ii","Z.Ii","Pr(z > 0)")
  ) %>%
  dplyr::mutate(
    coType = dplyr::case_when(
      `Pr(z > 0)` > 0.05 ~ "Insignificant",
      `Pr(z > 0)` <= 0.05 & Ii >= 0 & nta_trip_nodes$trip_count >= avg_trip_count ~ "HH",
      `Pr(z > 0)` <= 0.05 & Ii >= 0 & nta_trip_nodes$trip_count <  avg_trip_count ~ "LL",
      `Pr(z > 0)` <= 0.05 & Ii < 0 & nta_trip_nodes$trip_count >= avg_trip_count ~ "HL",
      `Pr(z > 0)` <= 0.05 & Ii < 0 & nta_trip_nodes$trip_count < avg_trip_count ~ "LH"
    )
  )

nta_trip_network_cluster <- nta_trip_network %>%
  tidygraph::activate(nodes) %>%
  dplyr::mutate(coType = lisa_classes$coType %>% tidyr::replace_na("Insignificant"))

nta_sig <- nta_trip_network_cluster %>%
  dplyr::filter(coType != "Insignificant")
```

Visualize the confidence type for significant nodes
```{r, message=FALSE}
ggraph::ggraph(nta_sig, layout="stress") +
  ggraph::geom_node_circle(aes(r = 0.025, color = coType))
```

### Natural log for trips
```{r, message=FALSE}
t_nta_trip_nodes <- nta_trip_nodes %>%
  dplyr::mutate(log_trip_count = log(trip_count))

t_local_moran <- spdep::localmoran(
  t_nta_trip_nodes$log_trip_count,
  nta_trip_network_weights,
  zero.policy = TRUE,
  na.action = na.omit,
)

sig_lev <- 0.05
t_avg_trip_count <- mean(t_nta_trip_nodes$log_trip_count)

t_lisa_classes <- t_local_moran %>%
  tibble::as_tibble() %>%
  magrittr::set_colnames(
    c("Ii","E.Ii","Var.Ii","Z.Ii","Pr(z > 0)")
  ) %>%
  dplyr::mutate(
    coType = dplyr::case_when(
      `Pr(z > 0)` > 0.05 ~ "Insignificant",
      `Pr(z > 0)` <= 0.05 & Ii >= 0 & t_nta_trip_nodes$log_trip_count >= t_avg_trip_count ~ "HH",
      `Pr(z > 0)` <= 0.05 & Ii >= 0 & t_nta_trip_nodes$log_trip_count <  t_avg_trip_count ~ "LL",
      `Pr(z > 0)` <= 0.05 & Ii < 0 & t_nta_trip_nodes$log_trip_count >= t_avg_trip_count ~ "HL",
      `Pr(z > 0)` <= 0.05 & Ii < 0 & t_nta_trip_nodes$log_trip_count < t_avg_trip_count ~ "LH"
    )
  )

t_nta_trip_network_cluster <- nta_trip_network %>%
  tidygraph::activate(nodes) %>%
  dplyr::mutate(coType = t_lisa_classes$coType %>% tidyr::replace_na("Insignificant"))

t_nta_sig <- t_nta_trip_network_cluster %>%
  dplyr::filter(coType != "Insignificant")
```

Visualize the confidence type for significant nodes
```{r, message=FALSE}
ggraph::ggraph(t_nta_sig, layout="stress") +
  ggraph::geom_node_circle(aes(r = 0.025, color = coType))
```
