---
output:
  pdf_document: default
  html_document: default
---
# Visualization of desire line network between NTAs

This notebook provides a reference visualization for a network of desire lines.
The nodes are the NTAs. They are labeled with the numeric portion of the NTA code.
The edges are the trips between them. The edges are weighted based on the number
of trips between the two NTAs. The graph is undirected. Consequently, a trip
in either direction is counted towards the weight for the edge. If there are no
trips between two NTAs, then they do not share an edge.

There is no statistical test performed on the network in this form. Instead, it
serves as foundation for the next step of analysis.

---
output:
  pdf_document: default
  html_document: default
---
```{r, message=FALSE}
library(tidyverse)
library(ggraph)
library(sf)
library(sfnetworks)
```

```{r, message=FALSE}
bk_name <- "Brooklyn"
bk_county_code <- "047"
bk_parks <- "BK99"
```

## Graph desire lines
Construct the graph of all desire lines between all Brooklyn NTAs


Places nodes and edges in data frame
```{r, message=FALSE}
nta_trips <- readr::read_csv('./data/nta-trip-network.csv')
nta_trip_nodes <- sf::st_read('./data/nyc_2010_nta_borders.geojson') %>%
  dplyr::filter(BoroName == bk_name)  %>%
  dplyr::filter(NTACode != bk_parks) %>%
  select("NTACode") %>%
  mutate(geometry = sf::st_point_on_surface(geometry))

nta_trip_edges <- tibble::tibble(from = nta_trips$nta_code_one, to = nta_trips$nta_code_two, weight = nta_trips$S000)
nta_trip_graph <- sfnetworks::sfnetwork(nodes = nta_trip_nodes, edges = nta_trip_edges, node_key = 'NTACODE')
```

Visualize the graph
```{r, message=FALSE}
ggraph::ggraph(nta_trip_graph, layout="stress") +
  geom_edge_link() +
  geom_node_circle(aes(r = 0.01), fill = "blue") +
  geom_node_text(aes(label = stringr::str_sub(NTACode, 3,4)), repel = TRUE)
```

## Subgraph 
The full graph is staturated with data. Here, I present subsections of the edges
to help visualize the structure of the full graph.

### Random Sample
A random sampling of edges 
```{r, message=FALSE}
nta_trip_edges_rand <- nta_trip_edges %>%
  dplyr::slice_sample(n = 50)
nta_trip_graph_rand <- sfnetworks::sfnetwork(nodes = nta_trip_nodes, edges = nta_trip_edges_rand, node_key = 'NTACode')
```

```{r, message=FALSE}
ggraph::ggraph(nta_trip_graph_rand, layout="kk") +
  geom_edge_link(mapping = aes(edge_width = weight)) +
  geom_node_circle(aes(r = 0.1), fill = "blue") +
  geom_node_text(aes(label = stringr::str_sub(NTACode, 3,4)), repel = TRUE)
```

### Most popular commutes
A sampling of the edges with the top 50 number of commutes
```{r, message=FALSE}
nta_trip_edges_top <- nta_trip_edges %>%
  dplyr::slice_max(order_by = weight, n = 50)
nta_trip_graph_top <- sfnetworks::sfnetwork(nodes = nta_trip_nodes, edges = nta_trip_edges_top, node_key = 'NTACode')
```

```{r, message=FALSE}
ggraph::ggraph(nta_trip_graph_top, layout="kk") +
  geom_edge_link(mapping = aes(edge_width = weight)) +
  geom_node_circle(aes(r = 0.2), fill = "blue") +
  geom_node_text(aes(label = stringr::str_sub(NTACode, 3,4)), repel = TRUE)
```