---
output:
  pdf_document: default
  html_document: default
---
# Network auto-correlation for desire lines between tracts in Brooklyn

The anaylsis paralles that of the NTA auto-correlation. However, it is at the
tract resolution, instead of the NTA resolution. Initially, I attempted to 
examine the auto-correlation for all tracts in Brooklyn. However, there are 133000
unique trips between tracts in Brooklyn. Checking whether an undirected edge between
each tract pair is 133000 choose 2, or close to 9billion operations. This overwhelmed
my computer. Instead, I arbitrarily picked an NTA and examined the trips between
tracts within that NTA. I repeated this a few times with different NTAs. Only one
analysis shown in these results. But, it would be straight-forward to adjust the code to
iterate through the NTAs and record the results.

```{r, message=FALSE}
library(tidyverse)
library(tidygraph)
library(ggraph)
library(spData)
library(spdep)
library(igraph)
options(scipen = 999)
```

Define each node as a trip 
```{r, message=FALSE}
bk_name <- "Brooklyn"
bk_county_code <- "047"

bk_ods <- readr::read_csv('./data/ny_od_main_JT00_2019.csv') %>%
  dplyr::filter(
    stringr::str_sub(as.character(w_geocode), 3, 5) == bk_county_code &
    stringr::str_sub(as.character(h_geocode), 3, 5) == bk_county_code
  )
```
*Note* Attempt to use all census tracts in borough overwhelmed my cpu
Code related to analyzing the whole borough is demoted to read-only

```
trip_nodes <- bk_ods %>%
  dplyr::mutate(
    w_tract = stringr::str_sub(as.character(w_geocode), 6, 11),
    h_tract = stringr::str_sub(as.character(h_geocode), 6, 11),
  ) %>%
  dplyr::filter(w_tract != h_tract) %>%
  dplyr::mutate(
    trip = stringr::str_c(
      "CT",
      ifelse(w_tract < h_tract, w_tract, h_tract),
      "CT",
      ifelse(w_tract > h_tract, w_tract, h_tract)
      )
  ) %>%
  dplyr::select(trip, S000) %>%
  dplyr::group_by(trip) %>%
  dplyr::summarise(
    S000 = sum(S000)
  ) %>%
  unique() 
```

Define each edge as a shared destination between two trips
```{r, message=FALSE}
build_edges <- function(nodes){
  edges_from <- vector()
  edges_to <- vector()
  nodes_count <- length(nodes)
  for(i in 1:(nodes_count - 1)) {
    offset <- i + 1
    from_node <- nodes[i]
    from_one <- stringr::str_sub(from_node, 1, 8)
    from_two <- stringr::str_sub(from_node, 9, 16)
    for(j in offset:nodes_count){
      to_node <- nodes[j]
      are_neighbors <- stringr::str_detect(to_node, from_one) | stringr::str_detect(to_node, from_two)
      if (are_neighbors) {
        edges_from <- append(edges_from, from_node)
        edges_to <- append(edges_to, to_node)
      }
    }
  }
  return (tibble::tibble(from = edges_from, to = edges_to))
}

```

```
trip_edges <- build_edges(trip_nodes$trip)

trip_network <- tidygraph::tbl_graph(
  nodes = trip_nodes,
  edges = trip_edges
  )

```
## Network auto-correlation of desire lines between census tracts within an NTA
```{r, message=FALSE}
tract_nta_equiv <- readxl::read_xlsx('./data/nyc_2010_census_tract_nta_equiv.xlsx')
bk_tract_nta_equiv <- tract_nta_equiv %>%
  dplyr::filter(borough_name == bk_name)

bk_nta_codes <- unique(bk_tract_nta_equiv$nta_code)

## Used the first number that popped into my head- psuedo random but repeatable
nta_of_interest <- bk_nta_codes[46]
tracts_oi <- bk_tract_nta_equiv %>%
  dplyr::filter(nta_code == nta_of_interest)
tracts_oi <- tracts_oi$census_tract

trips_nodes_oi <- bk_ods %>%
  dplyr::mutate(
    w_tract = stringr::str_sub(as.character(w_geocode), 6, 11),
    h_tract = stringr::str_sub(as.character(h_geocode), 6, 11),
  ) %>%
  dplyr::filter(w_tract != h_tract) %>%
  dplyr::filter(w_tract %in% tracts_oi & h_tract %in% tracts_oi) %>%
  dplyr::mutate(
    trip = stringr::str_c(
      "CT",
      ifelse(w_tract < h_tract, w_tract, h_tract),
      "CT",
      ifelse(w_tract > h_tract, w_tract, h_tract)
      )
  ) %>%
  dplyr::select(trip, S000) %>%
  dplyr::group_by(trip) %>%
  dplyr::summarise(
    S000 = sum(S000)
  ) %>%
  unique()
  
```

```{r, message=FALSE}
trip_edges_oi <- build_edges(trips_nodes_oi$trip)

trip_network_oi <- tidygraph::tbl_graph(
  nodes = trips_nodes_oi,
  edges = trip_edges_oi,
  node_key = "trip"
  )
```

```{r, message=FALSE}
total_trips_oi <- sum(trips_nodes_oi$S000)

ggraph::ggraph(trip_network_oi, layout="stress") +
  ggraph::geom_edge_link() +
  ggraph::geom_node_circle(aes(r = (trips_nodes_oi$S000 / total_trips_oi)), fill="blue") +
  ggraph::geom_node_text(aes(label = trips_nodes_oi$trip), repel = TRUE)
```
## Subgraph
Use a sub section of the network to make it easier to visualize

### Random sample
```{r, message=FALSE}
trip_nodes_rand <- trips_nodes_oi %>%
  dplyr::slice_sample(n = 50)
trip_edges_rand <- build_edges(trip_nodes_rand$trip)
trip_network_rand <- tidygraph::tbl_graph(
  nodes = trip_nodes_rand,
  edges = trip_edges_rand,
  node_key = "trip"
  )

total_trips_rand <- sum(trip_nodes_rand$S000)
ggraph::ggraph(trip_network_rand, layout="stress") +
  geom_edge_link() +
  geom_node_circle(aes(r = (trip_nodes_rand$S000 / total_trips_rand)), fill = "blue") +
  geom_node_text(aes(label = trip_nodes_rand$trip), repel = TRUE)
```

### Most popular trips
```{r, message=FALSE}
trip_nodes_rand <- trips_nodes_oi %>%
  dplyr::slice_max(order_by = S000, n = 50)
trip_edges_rand <- build_edges(trip_nodes_rand$trip)
trip_network_rand <- tidygraph::tbl_graph(
  nodes = trip_nodes_rand,
  edges = trip_edges_rand,
  node_key = "trip"
  )

total_trips_rand <- sum(trip_nodes_rand$S000)
ggraph::ggraph(trip_network_rand, layout="stress") +
  geom_edge_link() +
  geom_node_circle(aes(r = (trip_nodes_rand$S000 / total_trips_rand)), fill = "blue") +
  geom_node_text(aes(label = trip_nodes_rand$trip), repel = TRUE)
```

## Global Moran's I
```{r, message=FALSE}
trip_network_weights <- trip_network_oi %>%
  igraph::as_adj() %>%
  spdep::mat2listw()

global_morans <- spdep::moran.test(trips_nodes_oi$S000, trip_network_weights, zero.policy = TRUE)
global_morans  
```
```{r, message=FALSE}
spdep::moran.plot(
  trips_nodes_oi$S000,
  trip_network_weights,
  zero.policy = TRUE,
  xlab = "trip count",
  ylab = "lagged trip count",
  pch = 20,
)
```

## Local Indicators of spatial autocorrelation
```{r, message=FALSE}
local_moran <- spdep::localmoran(
  trips_nodes_oi$S000,
  trip_network_weights,
  zero.policy = TRUE,
  na.action = na.omit,
)

sig_lev <- 0.05
avg_trip_count <- mean(trips_nodes_oi$S000)

lisa_classes <- local_moran %>%
  tibble::as_tibble() %>%
  magrittr::set_colnames(
    c("Ii","E.Ii","Var.Ii","Z.Ii","Pr(z > 0)")
  ) %>%
  dplyr::mutate(
    coType = dplyr::case_when(
      `Pr(z > 0)` > 0.05 ~ "Insignificant",
      `Pr(z > 0)` <= 0.05 & Ii >= 0 & trips_nodes_oi$S000 >= avg_trip_count ~ "HH",
      `Pr(z > 0)` <= 0.05 & Ii >= 0 & trips_nodes_oi$S000 <  avg_trip_count ~ "LL",
      `Pr(z > 0)` <= 0.05 & Ii < 0 & trips_nodes_oi$S000 >= avg_trip_count ~ "HL",
      `Pr(z > 0)` <= 0.05 & Ii < 0 & trips_nodes_oi$S000 < avg_trip_count ~ "LH"
    )
  )

trip_network_cluster <- trip_network_oi %>%
  tidygraph::activate(nodes) %>%
  dplyr::mutate(coType = lisa_classes$coType %>% tidyr::replace_na("Insignificant"))

tract_sig <- trip_network_cluster %>%
  dplyr::filter(coType != "Insignificant")

ggraph::ggraph(tract_sig, layout="stress") +
  ggraph::geom_node_circle(aes(r = 0.025, color = coType))

```
